% arara: pdflatex: { synctex: yes }
% arara: makeindex: { style: ctuthesis }
%% arara: bibtex


%\listfiles


%\PassOptionsToPackage{cp1250}{inputenc}

% The class takes all the key=value arguments that \ctusetup does,
% and couple more: draft and oneside
\documentclass[twoside]{ctuthesis}

\usepackage{graphicx}
\usepackage{svg}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{textgreek}
\usepackage{pdfpages}
\usepackage{bytefield}
\usepackage{tikz-timing}
\usepackage{soul}

\usepackage{float}
\usepackage[defernumbers=true, sorting=none]{biblatex}
\addbibresource{bibliography.bib}


\newcolumntype{?}{!{\vrule width 2pt}}

\usepackage{svg}
\svgsetup{
    inkscape=pdf,
    inkscapeexe={C:/inkscape.exe.lnk -z -C"}
}
\graphicspath{{images/}}

\usepackage[per-mode=symbol,separate-uncertainty=true]{siunitx}


\DeclareSIUnit\bit{b}


\makeatletter
\edef\mytoday{\expandafter\@gobbletwo\the\year\ifnum\month<10 0\fi\the\month\ifnum\day<10 0\fi\the\day}
\makeatother

% LaTeX logo with better kerning in sf bf font
\makeatletter
\newcommand\LaTeX@lmss@bx{L\kern -.33em{\sbox \z@ T\vbox to\ht \z@ {\hbox {\check@mathfonts \fontsize \sf@size \z@ \math@fontsfalse \selectfont A}\vss }}\kern -.15em\TeX}
\DeclareRobustCommand\myLaTeX{%
	\ifcsname LaTeX@\f@family @\f@series\endcsname
		\csname LaTeX@\f@family @\f@series\endcsname
	\else
		\LaTeX
	\fi
}

\ctusetup{
%	preprint = {\ctuverlog \\ ctuman \mytoday},
	mainlanguage = english,
	titlelanguage = english,
	otherlanguages = {english},
	title-czech = {Vyčítací systém pro FastIC+},
	title-english = {FastIC+ readout system},
	doctype-english = {Diploma Thesis},
	xfaculty = F3,
        xdoctype = M,
	department-english = {Department of Microelectronics},
%	department-english = {Department of Mathematics},
	author = {Bc. Vojtěch Vosáhlo},
	supervisor = {Ing. Vít Záhlava, CSc.},
        fieldofstudy-english = {Electronics},
%	supervisor-address = {Ústav X, \\ Uliční 5, \\ Praha 99},
	keywords-czech = {FastIC+, Vyčítací systém, Rychlé časování, CERN, DPS},
	keywords-english = {FastIC+, Readout system, Fast Timing, CERN, PCB},
	day = 25,
	month = 5,
	year = 2025,
%	list-of-figures = false,
%	list-of-tables = false,
%	monochrome = true,
%	savetoner = true,
	pkg-listings = true,
        pkg-makeidx = true,
	ctulstbg = none,
%	layout-short = true,
}

\ctuprocess

% Theorem declarations, this is the reasonable default, anybody can do what they wish.
% If you prefer theorems in italics rather than slanted, use \theoremstyle{plainit}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{note}
\newtheorem*{remark*}{Remark}
\newtheorem{remark}[theorem]{Remark}

% Marginpars used as navigation aids.
\usepackage{mparhack}
\usepackage{makecell}

\newcommand\indexmp[1]{{\sffamily\bfseries#1}}

\ExplSyntaxOn
\cs_new:Nn \ctuman_domarginpar:n {
	\marginpar
	[ \raggedleft \footnotesize \sffamily #1 ]
	{ \raggedright \footnotesize \sffamily #1 }
}
\cs_generate_variant:Nn \ctuman_domarginpar:n { x }
\DeclareDocumentCommand \ctump { m } {
	\clist_set:Nn \ctuman_temp_clist { #1 }
	\ctuman_domarginpar:x { \clist_use:Nnnn \ctuman_temp_clist { \\ } { \\ } { \\ } }
	\clist_map_inline:Nn \ctuman_temp_clist { \index{##1|indexmp} }
	\ignorespaces
}
\ExplSyntaxOff

\sisetup{locale = UK}

% Abstract in Czech
\begin{abstract-czech}
Tato práce se zabývá návrhem elektronického systému určeného k vyčítání dat z integrovaných obvodů FastIC+ vyvíjených v rámci CERN pro aplikace precizního rychlého časování. Práce nastiňue vlastnosti zmíněných ASIC, jejich účel a popisuje komunikaci s čipy. Následně detailně popisuje návrh elektroniky, firmware zařízení a jejich testování. Dále popisuje návrh ovládacího software pro měření se zařízením a interpretaci dat. V neposlední řadě pak ověřuje funkcionalitu zařízení na praktickém experimentu. 
\end{abstract-czech}

% Abstract in English
\begin{abstract-english}
This thesis focuses on the design of an electronic system intended for data readout from FastIC+ integrated circuits developed within CERN for precise fast timing applications. The thesis outlines the properties of the mentioned ASICs, their purpose, and describes communication with the chips. Subsequently, it provides a detailed description of the electronics design, firmware development, and final testing of the device. Furthermore, it describes the development of control software for measurements with the device and data interpretation. Finally, it verifies the functionality of the device in a practical experiment.
\end{abstract-english}

% Acknowledgements / Podekovani
\begin{thanks}
----
\end{thanks}

% Declaration / Prohlaseni
\begin{declaration}

I declare that this work is all my own work and I have cited all sources I have
used in the bibliography.
\medskip

Prague,  \ctufield{day}.~\monthinlanguage{title}~\ctufield{year}

\vspace*{2cm}

Prohlašuji, že jsem předloženou práci vypracoval samostatně, a že jsem uvedl veškerou použitou literaturu.

\medskip

V Praze, \monthinlanguage{second} \ctufield{day}, \ctufield{year}
\end{declaration}

\usepackage{url}

\usepackage{tabularx,array}

\usepackage{mathtools,amssymb}
\usepackage{threeparttable}[para,online,flushleft]
\usepackage{makecell}


\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}

% A savebox for typesetting listings in the titles
\newsavebox{\myboxa}

%\newcommand*\symbO{$\color{red}\bowtie$}
\newcommand*\symbO{\raisebox{0.5\height}{\scalebox{0.7}{\color{red}${\vartriangleright}\mkern-6mu{\vartriangleleft}$}}}
\newcommand*\symbM{\raisebox{0.5\height}{\scalebox{0.7}{\color{red}${\blacktriangleright}\mkern-6mu{\blacktriangleleft}$}}}
\newcommand*\itemO{\item\leavevmode\kern-0.33em\symbO}
\newcommand*\itemM{\item\leavevmode\kern-0.33em\symbM}



\begin{document}

% We actually don't want inline listings to have a background color
\renewcommand \ctulstsep {0pt}

% \ctuclsname for typesetting the class' name
\newcommand\ctuclsname{\leavevmode\unhcopy\ctuclsnamebox}
\newsavebox\ctuclsnamebox
\begin{lrbox}{\ctuclsnamebox}
\ctulst!ctuthesis!
\end{lrbox}

\maketitle


\newpage
\begingroup
\color{black}
\endgroup

\pagebreak

\chapter{Introduction}
The development of devices for particle detectors at CERN has often found applications beyond high-energy physics, particularly in the medical field. One such example is the FastIC chip, which was designed for fast timing applications but has been also utilized in positron emission tomography (\emph{PET}) to improve imaging precision. Building on the success of the FastIC chip, a new version, FastIC+, has been developed. This updated version integrates a picosecond Time-to-Digital Converter (\emph{TDC}) for enhanced time measurement capabilities and higher integration. To facilitate the evaluation of the FastIC+ chip's performance, a evaluation board had to be designed to provide a compact and accessible platform for researchers and engineers to test and analyze the chip's capabilities in various applications.

This work describes the design of the evaluation board, referred to as the readout system, developed at CERN. First, the FastIC+ is detailed along with its communication protocol. Subsequently, the hardware concept is proposed, and its key components are described in detail. The hardware design process is presented, and both firmware and host software are developed to enable easy control of the device. Finally, functional tests are conducted to evaluate the device, and an experiment at CERN is performed to verify the functionality of the readout system in a real-world application.

\include{fastic}
\include{aurora}
\include{fastic_com}
\include{electronics}
\include{mechanical}
\include{firmware}
\include{software}
\include{tests}



\clearpage
\FloatBarrier
\printbibliography

\chapter*{List of abbreviations}
\noindent
\begin{tabularx}{\linewidth}
{ l >{\raggedright\arraybackslash}X }
\bfseries Abbreviation & \bfseries Explanation\\\Midrule
CERN & Conseil Européen pour la Recherche Nucléaire \\
ASIC & Application Specific Integrated Circuit \\
PET & Positron Emission Tomography\\
TDC & Time-to-Digital Converter\\
SiPM & Silicon Photo Multiplier\\
PMT & Photo Multiplier Tube\\
MCP & Micro Channel Plate\\
ToA & Time of Arrival\\
ToT & Time over Threshold\\
ToF PET & Time of Flight Positron Emission Tomography\\
LIDAR & Light Detection and Ranging\\
I2C & Inter-Integrated Circuit\\
FSM & Finite State Machine\\
FERO & Front-End Readout\\
BERO & Back-End Readout\\
FIFO & First In First Out\\
MUX & Multiplexer\\
SLVS & Scalable Low Voltage Signaling\\
MSB & Most Significant Bit\\
LSB & Least Significant Bit\\
BTF & Block Type Field\\
USB & Universal Serial Bus\\
FPGA & Field Programmable Gate Array\\
HDL & Hardware Description Language\\
IP cores & Intellectual Property Cores\\
ADC & Analog to Digital Converter\\
DAC & Digital to Analog Converter\\
SPI & Serial Peripheral Interface\\
DMA & Direct Memory Access\\
TFBGA & Thin Fine Ball Grid Array\\
MCU & Microcontroller Unit\\
PLL & Phase Locked Loop\\
LVDS & Low Voltage Differential Signaling\\
CMOS & Complementary Metal-Oxide-Semiconductor\\
EMI & Electromagnetic Interference\\
SMA & SubMiniature version A\\
ESD & Electrostatic Discharge\\
PHY & Physical Layer\\
ULPI & UTMI+ Low Pin Interface\\
UTMI & USB Transceiver Macrocell Interface\\
UCPD & USB Type-C Power Delivery\\
RMS & Root Mean Square\\
LDO & Low Dropout Regulator\\
EEPROM & Electrically Erasable Programmable Read-Only Memory\\
THT & Through-Hole Technology\\
PCB & Printed Circuit Board\\
\end{tabularx}

\end{document}